---
title: "Project 2"
output:
  pdf_document: default
  html_document: default
---
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

library(DBI) 
library(dplyr) 
library(lubridate)
library(tidyverse)
library(janitor)
library(dplyr)
library(tidyr)
library(stringr)
library(httr)
library(purrr)
library(magrittr)
library(knitr)

politicians_db <- dbConnect(RSQLite::SQLite(), "zh_politicians.db") 
dbListTables(politicians_db)
mandates_tbl <- tbl(politicians_db, "MANDATES") %>% as_tibble() %>% clean_names 
addresses_tbl<-tbl(politicians_db, "ADDRESSES") %>% as_tibble() %>% clean_names
persons_tbl<-tbl(politicians_db, "PERSONS") %>% as_tibble() %>% clean_names
affiliations_tbl<-tbl(politicians_db, "AFFILIATIONS") %>% as_tibble() %>% clean_names

```

# Part 1

Number of people with an active mandate changed over the years

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

MD_tbl_short<-mandates_tbl %>% 
  select(-mandate_start_day,-mandate_start_month,
         -mandate_end_day,-mandate_end_month)
AF_tbl_short<-affiliations_tbl%>% 
  select(-affiliation_start_day,-affiliation_start_month,
         -affiliation_end_day,-affiliation_end_month)

mandate_affiliation_left_tbl<-left_join(MD_tbl_short,AF_tbl_short, by=c("id"))

mandates_counted_years<-mandate_affiliation_left_tbl %>% 
  filter(mandate_start_year>0,mandate_end_year>0) %>% 
  mutate(active_year=map2(mandate_start_year,mandate_end_year,seq)) %>% 
  group_by(assembly) %>% 
  unnest_longer(col = active_year) %>% 
  count(active_year) %>%
  ungroup()
  
mandates_counted_years %>% drop_na() %>% 
  ggplot(mapping=aes(x=active_year, 
                     y=n)) + xlim(1800, 2020)+
  geom_line(aes(group=assembly, color=assembly)) +
  theme(legend.position="right", legend.box = "horizontal")+
  labs (title = "Number of active mandates each year",
        subtitle = "Using colors to differentiate assembly categories",
        caption = "Source: Politicians data",
        x = "Year",
        y = "Total number of active mandates")

```

# Part 2

Facet charts with one chart per assembly seperated for men and women

```{r message=FALSE, warning=FALSE}

mandates_counted_years_assembly_gender<-mandate_affiliation_left_tbl %>% 
  left_join(persons_tbl, by=c("id"))%>% 
  filter(mandate_start_year>0,mandate_end_year>0, !is.na(gender)) %>% 
  mutate(active_year=map2(mandate_start_year,mandate_end_year,seq)) %>% 
  group_by(assembly,gender) %>% 
  unnest_longer(col = active_year) %>% 
  count(active_year) %>%
  ungroup()

mandates_counted_years_assembly_gender %>% 
  ggplot(mapping=aes(x=active_year, 
                     y=n)) + xlim(1800, 2020)+
  geom_line(aes(group=assembly, color=assembly)) +
  facet_grid(vars(gender)) +
  theme(legend.position="right", legend.box = "horizontal")+
  labs (title = "Number of active mandates each year",
        subtitle = "Using colors to differentiate assembly categories and facet grid for men and women",
        caption = "Source: Politicians data",
        x = "Year",
        y = "Total number of active mandates")
```

# Part 3

Proportion of elected politicians from each party in year 2000

```{r message=FALSE, warning=FALSE}

Elected_politician_2000<- mandate_affiliation_left_tbl%>%
  filter(affiliation_start_year>0,affiliation_end_year>0 & affiliation_end_year<2020) %>% 
  select(id,assembly,party,affiliation_start_year,affiliation_end_year)%>%
  mutate(affiliation_active_year=map2(affiliation_start_year,affiliation_end_year,seq)) %>% 
  group_by(party) %>% 
  unnest_longer(col = affiliation_active_year) %>% 
  filter(affiliation_active_year==2000)%>%
  ungroup()

Proportion_Elected_politician_2000<-Elected_politician_2000%>% 
  group_by(assembly,party)%>%
  summarise(n=n())%>%
  mutate(proportion=n/sum(n))%>%
  ungroup(party)

Table<-knitr::kable(Proportion_Elected_politician_2000, caption = "Table with kable")
Table

ggplot(Proportion_Elected_politician_2000, aes(x="", y=proportion, fill=party)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +  facet_grid(facets=. ~ assembly)+
  theme_void() +
  labs (title = "Number of active mandates each year",
        subtitle = "facets to differentiate assemblies",
        caption = "Source: Politicians data")
```

# Part 4 

Change in the composition of assemblies over the years

```{r message=FALSE, warning=FALSE}

Composition_assembly_party_by_year<- mandate_affiliation_left_tbl%>%
  filter(affiliation_start_year>0,affiliation_end_year>0 & affiliation_end_year<2020) %>% 
  select(id,assembly,party,affiliation_start_year,affiliation_end_year)%>%
  mutate(affiliation_active_year=map2(affiliation_start_year,affiliation_end_year,seq)) %>% 
  unnest_longer(col = affiliation_active_year) %>% 
  group_by(assembly,affiliation_active_year, party) %>% 
  summarise(n=n())%>%
  mutate(proportion=n/sum(n))%>%
  ungroup(affiliation_active_year, party)

Composition_assembly_party_by_year %>% 
  ggplot(mapping=aes(x=affiliation_active_year, 
                     y=proportion)) + 
  geom_line(aes(group=party, color=party)) +
  facet_grid(vars(assembly)) +
  theme(legend.position="none")+
  labs (title = "composition of assemblies across year",
        subtitle = "Using colors to differentiate party and facet grid for assemblies",
        caption = "Source: Politicians data",
        x = "Year",
        y = "Proportion of party in an assembly")

# Problem in placing legend box on right. The whole screen is covered by the legend box

```

# Part 5

Polticians with year of death in the data

```{r message=FALSE, warning=FALSE}

Politicians_death<-persons_tbl%>%
  arrange(desc(year_of_death))%>%
  filter(year_of_death>=0,year_of_birth!=0) %>%
  mutate(lifespan=as.numeric(year_of_death)-as.numeric(year_of_birth))

Average_politicians_age<-Politicians_death%>%
  select(title, lifespan)%>%
  mutate(title = na_if(title, ""))%>%
  mutate(is_title=ifelse(is.na(title),"no_title","with_title"))%>%
  group_by(is_title)%>%
  summarise(average_lifespan=mean(lifespan))%>%
  ungroup()
Politicians_death_table <- Politicians_death %>%
  select(title, lifespan)%>%
  mutate(title = na_if(title, ""))%>%
  mutate(is_title=ifelse(is.na(title),"no_title","with_title")) %>% 
  select(is_title, lifespan)
Politicians_death_table %>%  
  ggplot2::ggplot(mapping = aes( y = lifespan, group=is_title)) +
  geom_boxplot(aes(colour=is_title))

Politicians_death_table %>% 
  select(is_title,lifespan) %>% 
  infer::t_test( lifespan ~ is_title,
                 order = c("no_title","with_title"),
                 var.equal = FALSE)
  
```

p-value looks quite significant for comparison of average lifespan for people with and wihout title.

# Part 6 

Polticians born before and after 1918

```{r message=FALSE, warning=FALSE}

Politicians_born_1918_title<-Politicians_death%>%
  arrange(desc(year_of_birth))%>%
  filter(year_of_birth!=0) %>% 
  select(id, year_of_birth, title, lifespan) %>% 
  mutate(title = na_if(title, ""))%>%
  mutate(is_title=ifelse(is.na(title),"no_title","with_title")) %>% 
  mutate(is_born=ifelse(year_of_birth<1918,"before_1918","after_1918"))%>% 
  select(id, is_title, is_born, lifespan)

Politician_1918_avg_lifespan<-Politicians_born_1918_title %>%
  group_by(is_born,is_title)%>%
  summarise(average_lifespan=mean(lifespan))

# Filter data sets to make two data sets for t-test

Politician_before_1918<-Politicians_born_1918_title %>%
  filter(is_born=="before_1918")

Politician_before_1918 %>%  
  ggplot2::ggplot(mapping = aes( x=is_title,y = lifespan, group=is_title)) +
  geom_boxplot(aes(colour=is_title)) + 
  labs(title = "Politicians before 1918")

Politician_before_1918 %>% select(is_title,lifespan) %>% 
  infer::t_test( lifespan ~ is_title,
                 order = c("no_title","with_title"),
                 var.equal = TRUE)

Politician_after_1918<-Politician_before_1918<-Politicians_born_1918_title %>%
  filter(is_born=="after_1918")
Politician_after_1918 %>%  
  ggplot2::ggplot(mapping = aes( x=is_title,y = lifespan, group=is_title)) +
  geom_boxplot(aes(colour=is_title))+ 
  labs(title = "Politicians after 1918")

Politician_after_1918 %>% select(is_title,lifespan) %>% 
  infer::t_test( lifespan ~ is_title,
                 order = c("no_title","with_title"),
                 var.equal = TRUE)
 

```
 There is difference in lifespan for politicians borne before 1918 with and without title.
 
 p-value looks quite significant for comparison of average lifespan for politician born before 1918 with and wihout title.

p-value does not look significant for comparison of average lifespan for politician born after 1918 with and wihout title.

# Part 7

Most number of mandates by a politician

```{r message=FALSE, warning=FALSE}

number_mandates_by_politician<- mandate_affiliation_left_tbl%>%
  drop_na() %>% 
  group_by(person_id.x)%>%
  summarise(number_mandates=n())%>%
  ungroup%>%
  top_n(10)
Table<-knitr::kable(number_mandates_by_politician, caption = "Table with kable")

number_mandates_by_politician %>% 
  ggplot(aes(x=as.character(person_id.x), y=number_mandates))+
  geom_bar(stat="identity")+ 
  labs(title = "Number of mandates by a poltiician")
 
```

# Part 8

Politicians associated with multiple mandates at the same time

```{r message=FALSE, warning=FALSE}

politicians_multiple_mandates<-mandate_affiliation_left_tbl %>% 
  filter(mandate_start_year>0,mandate_end_year>0) %>% 
  mutate(active_year=map2(mandate_start_year,mandate_end_year,seq)) %>% 
  group_by(person_id.x) %>% 
  unnest_longer(col = active_year) %>% 
  count(active_year)%>%
  ungroup()

count_politicians_multiple_mandates<-politicians_multiple_mandates %>% 
  group_by(n,active_year) %>% 
  count("politician with x number of mandates") %>%
  ungroup(active_year)

count_politicians_multiple_mandates %>% 
  ggplot(mapping=aes(x=active_year, 
                     y=nn)) + 
  geom_line(aes(group=nn, color=nn)) +
  facet_grid(vars(n)) +
  theme(legend.position="none")+
  labs (title = "composition of politiciany by x number of mandates",
        subtitle = "Using colors to differentiate x number of mandates",
        caption = "Source: Politicians data",
        x = "Year",
        y = "Total number of policians with x number of mandates for that active mandate year")
```

# Part 9

Politicians associated with different parties

```{r message=FALSE, warning=FALSE}

Policitian_party<-mandate_affiliation_left_tbl%>%
  select(id, person_id.x,party)%>%
  group_by(person_id.x)%>%
  mutate(number_of_party=n_distinct(party)) %>% 
  arrange(desc(number_of_party)) %>% 
  ungroup() %>% 
  top_n(4)


Policitian_party_table<-Policitian_party %>% 
  kable()

```

Politicians associated with different parties are: `r Policitian_party_table`.

# Part 10 

Leaflet and ggmap for addresses of politicians

```{r}

library(dplyr)
library(purrr)
library(magrittr)

politician_address<-addresses_tbl%>%
                  select(id, person_id,street, house_number, city, postal_code)%>%
                  mutate_all(na_if,"") %>% 
                  dplyr::filter(complete.cases(.)) 


politician_address_20<-politician_address %>% 
                    mutate(full_address=str_c(street, house_number, postal_code, city, sep=",")) %>% 
                    mutate(postal_code=as.numeric(str_extract(politician_address$postal_code, "\\d+"))) %>% 
                    mutate(house_number=as.numeric(str_extract(politician_address$house_number, "\\d+"))) %>%
                    mutate(street=str_extract(politician_address$street, "\\w+")) %>%
                    mutate(country="switzerland") %>% 
                    mutate_all(na_if,"") %>% 
                    dplyr::filter(complete.cases(.)) %>% 
                    mutate(address_geocode=str_glue_data(politician_address,"{street}","{house_number}","{postal_code}","{city}","{country}", .sep="+")) %>% 
                    top_n(3)

slowGET <- function( url ) {
  Sys.sleep(0.5)
  print( url )
  return ( GET(url) )
}

latitude_longitude <-str_glue("https://geocode.xyz/",
                              "{politician_address_20$address_geocode}",
                              "?json=1") %>% 
  map(GET) %>%
  keep(~ status_code(.) == 200) %>% 
  map(content)
  

latitude_longitude_extract<-latitude_longitude %>%
 purrr::map_df(magrittr::extract, c("latt", "longt")) %>% 
 mutate(latt=as.numeric(latt),longt=as.numeric(longt))

# I have commented the leaflet code below

#library(leaflet)
#latitude_longitude_extract %>%
#leaflet() %>% 
#addTiles() %>% 
#addMarkers(lng=~longt,lat=~latt)

library(ggmap)
switzerland_area <- c(left=5.91, bottom=45.89, right=10.54, top=47.84)
switzerland_map <- get_stamenmap(bbox=switzerland_area, zoom=7)
ggmap(switzerland_map) +
  geom_point(data=latitude_longitude_extract , 
             aes(x=longt, y=latt, alpha=.5))+
  labs(title="Where are policiticians in Switzerland?",
       subtitle="Geolocation of 20 listings from the politician data in Switzerland")
```
