---
title: "R Notebook"
output: html_document
---

```{r echo=TRUE, message=FALSE, warning=FALSE}

library(stringr)
library(httr)
library(purrr)
library(tidyr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(knitr)

```

# Part 1

Getting access to "The Movie DB"

```{r echo=TRUE, message=FALSE, warning=FALSE}

highest_gross_drama_2010 <-str_glue("https://api.themoviedb.org/3/",
                                    "discover/movie",
                                    "?api_key=55b3238346b11e88d118b5a647963747",
                                    "&with_genres=18&primary_release_year=2010&sort_by=revenue.desc") %>% 
  GET () %>% 
  content () %>%
  pluck("results") %>% 
  purrr::map_df(magrittr::extract,c ("title")) %>% 
  kable()

                             
actors_together<- str_glue("https://api.themoviedb.org/3/",
                           "discover/movie",
                           "?api_key=55b3238346b11e88d118b5a647963747",
                           "&with_cast=23659&with_cast=3896&sort_by=popularity.desc") %>% 
  GET() %>% 
  content() %>%
  pluck("results") %>% 
  #map(pluck,"title") %>% 
  purrr::map_df(magrittr::extract,c ("title")) %>% 
  kable()
  #map_df(extract, c("title"))  Not sure why map_df does not work

tom_cruise_kidsmovies<- str_glue("https://api.themoviedb.org/3/",
                                 "discover/movie",
                                 "?api_key=55b3238346b11e88d118b5a647963747",
                                 "&certification.lte=G&with_cast=500") %>% 
  GET() %>% 
  content() %>%
  pluck("results") %>%   
  purrr::map_df(magrittr::extract,c ("title")) %>% 
  kable() 
# map_df(extract, c("title"))
```

The highest grossing dramas from 2010 are :`r highest_gross_drama_2010`

Movie with actors Ferrell and Liam Neeson are : `r actors_together`

The kids movie done by Tom cruise are : `r tom_cruise_kidsmovies`

# Part 2

Extract id of the Pixar company

```{r echo=TRUE, message=FALSE, warning=FALSE}


Pixar <- str_glue(
  "https://api.themoviedb.org/3/",
  "search/company",
  "?api_key=55b3238346b11e88d118b5a647963747",
  "&query=Pixar&page=1"
) %>%
  GET %>%
  content %>%
  pluck("results") %>%
  purrr::map_df(magrittr::extract, c ("id", "name"))

Pixar_id_found<-Pixar %>% 
  filter(name=="Pixar") %>% 
  kable()
  

Pixar_revenue_movies <- str_glue(
  "https://api.themoviedb.org/3/",
  "discover/movie",
  "?api_key=55b3238346b11e88d118b5a647963747",
  "&sort_by=revenue.desc&include_adult=false&include_video=false&page=1&with_companies=3"
) %>%
  GET() %>%
  content() %>%
  pluck("results") %>%
  purrr::map_df(magrittr::extract, c ("original_title")) %>%
  kable() 

```

Pixar id was found to be `r Pixar_id_found`.

Pixar  movies sorted by descending revenue are :`r Pixar_revenue_movies`

# Part 3
 
Pixar movies sorted by descending revenue
 
```{r echo=TRUE, message=FALSE, warning=FALSE}

pages<-as.character(seq(1,6,1))

Pixar_revenue_movies_id <-str_glue("https://api.themoviedb.org/3/",
                                   "discover/movie",
                                   "?api_key=55b3238346b11e88d118b5a647963747",
                                   "&sort_by=revenue.desc&include_adult=false&include_video=false&page={pages}&with_companies=3") %>% 
  map(GET) %>% 
  map(content)%>%
  map(pluck, "results") %>% 
  flatten() %>% 
  map_df(magrittr::extract, c("original_title", "id")) %>% 
  kable()

```

Pixar  movies sorted by descending revenue are :`r Pixar_revenue_movies_id`

# Part 4

Comparison of the popularity of both Disney and Pixar films that came out from 2006 onwards.

```{r echo=TRUE, message=FALSE, warning=FALSE}

 # Part 4
 
Disney<- str_glue("https://api.themoviedb.org/3/",
                  "search/company",
                  "?api_key=55b3238346b11e88d118b5a647963747",
                  "&query=Disney&page=1") %>% 
  GET() %>% 
  content() %>%
  pluck("results") %>% 
  map_df(magrittr::extract, c("name", "id")) 

Disney_id_table<-Disney%>% 
  kable()

Disney_movies<- str_glue("https://api.themoviedb.org/3/",
                         "discover/movie",
                         "?api_key=55b3238346b11e88d118b5a647963747",
                         "&include_adult=false&include_video=false&page={pages}&primary_release_year=>%3D2006&with_companies=2") %>% 
  map(GET) %>% 
  map(content)%>%
  map(pluck, "results") %>% 
  flatten() %>% 
  map_df(magrittr::extract, c("original_title","release_date","vote_average", "vote_count","popularity"))

Disney_movie_table<-Disney_movies %>% 
  kable()

Pixar_movies <- str_glue(
  "https://api.themoviedb.org/3/",
  "discover/movie",
  "?api_key=55b3238346b11e88d118b5a647963747",
  "&include_adult=false&include_video=false&page={pages}&primary_release_year=>%3D2006&with_companies=3"
) %>%
  map(GET) %>%
  map(content) %>%
  map(pluck, "results") %>%
  flatten() %>%
  map_df(
    magrittr::extract,
    c("original_title","release_date", "vote_average","vote_count","popularity")
  ) 

Pixar_movie_table<-Pixar_movies %>% 
  kable()

Pixar_filtered<-Pixar_movies %>%
                mutate(company_name="Pixar") %>% 
                filter(vote_count>=50)
Disney_filtered<-Disney_movies %>%
                mutate(company_name="Disney") %>% 
                filter(vote_count>=50)
Disney_Pixar<-bind_rows(Pixar_filtered,Disney_filtered)


Disney_Pixar %>%
  select(vote_average,popularity,company_name) %>% 
  ggplot2::ggplot(mapping = aes(x = company_name, y = vote_average, group=company_name)) +
  geom_boxplot(aes(colour=company_name))

Disney_Pixar %>% 
  select(vote_average,popularity,company_name) %>%
  ggplot(mapping = aes(x = vote_average, fill = company_name)) +
  geom_histogram(binwidth = 2)

Disney_Pixar %>% 
  select(vote_average,popularity,company_name) %>% 
  infer::t_test(vote_average ~ company_name,
         order = c("Pixar", "Disney")
         ,var.equal = FALSE)

```
Searching for Disney id :`r Disney_id_table`

Disney id (Walt Disney Pictures) was found to be 2.

vote averages and vote counts for Disney movies :`r Disney_movie_table`

vote averages and vote counts for Pixar movies :`r Pixar_movie_table`

Using the box plot, it looks __vote_average__  is comparable for this set of movies (from year 2006) from Pixar and Disney.

Here p-value is greater than 0.05  which means there is not a strong relationship between the variables __vote_average__ and __company_name__ of movies (from year 2006) from Pixar and Disney.
