---
title: 'Project1: Time Series'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

library(readr)
library(tidyverse)
library(dplyr)
library(janitor)
library(kableExtra)
library(tidyr)
library(slider)

```

# Part1: Import the data

```{r}

cheese_data <- read_csv("Data/cheese_data.csv")
cheese_data_f1<-cheese_data %>% 
  filter(factory=="f1")
cheese_data_f2<-cheese_data %>% 
  filter(factory=="f2")
```

# Part2: Clean the data (manage missing values)

```{r}

cheese_data_f1_filtered<-cheese_data_f1 %>% 
  mutate_all(na_if, 0)

cheese_data_f2_filtered<-cheese_data_f2 %>% 
  mutate_all(na_if, 0)
  
count_missing_f1<-sum(is.na(cheese_data_f1_filtered))
  
count_missing_f2<-sum(is.na(cheese_data_f2_filtered))
  
dQ1_f1<-cheese_data_f1_filtered %>% 
    mutate(across(m1:m5,is.na)) %>% 
    mutate(total_na=m1+m2+m3+m4+m5) %>% 
    filter(total_na>1) %>% 
    select(factory, week, timepoint)
  
dQ2_f1<-cheese_data_f1_filtered %>% 
    anti_join(dQ1_f1)
  
dQ2_replaced_f1<-dQ2_f1 %>% 
    pivot_longer(cols = m1:m5, names_to ="measurements", values_to ="value") %>% 
    group_by(timepoint) %>% 
    mutate(four_timepoint_mean = sum(value, na.rm = TRUE)/4) %>%
    mutate(value=if_else(is.na(value), four_timepoint_mean, value)) %>% 
    select(-four_timepoint_mean)

dQ1_f2<-cheese_data_f2_filtered %>% 
    mutate(across(m1:m5,is.na)) %>% 
    mutate(total_na=m1+m2+m3+m4+m5) %>% 
    filter(total_na>1) %>% 
    select(factory, week, timepoint)
  
dQ2_f2<-cheese_data_f2_filtered %>% 
    anti_join(dQ1_f1)
  
dQ2_replaced_f2<-dQ2_f2 %>% 
    pivot_longer(cols = m1:m5, names_to ="measurements", values_to ="value") %>% 
    group_by(timepoint) %>% 
    mutate(four_timepoint_mean = sum(value, na.rm = TRUE)/4) %>%
    mutate(value=if_else(is.na(value), four_timepoint_mean, value)) %>% 
    select(-four_timepoint_mean)

count_timepointmissing_f1<-dQ1_f1 %>% 
  count()

count_timepointmissing_f2<-dQ1_f2 %>% 
  count()
  
```

There are `r count_missing_f1` missing values for data in factory f1 and `r count_missing_f2` missing values for data in factory f2.

There are `r count_timepointmissing_f1` missing timepoint values for data in factory f1 and `r count_timepointmissing_f2` missing timepoint values for data in factory f2.

# Part3: Manage unexpected values

```{r}
  
dQ3_expected_f1<-dQ2_replaced_f1 %>%
  filter(between(value, 15, 65))

dQ3_f1_unexpected<-dQ2_replaced_f1 %>% 
    anti_join(dQ3_expected_f1) %>% 
    nrow()
  

dQ3_expected_f2<-dQ2_replaced_f2 %>%
  filter(between(value, 15, 65))

dQ3_f2_unexpected<-dQ2_replaced_f2 %>% 
    anti_join(dQ3_expected_f2) %>% 
  nrow()
  

```
  
The unexpected or suspicious values are  `r dQ3_f1_unexpected` unexpected values for data in factory f1 and `r dQ3_f2_unexpected` unexpected values for data in factory f2.  

# Part4: Compute monitoring statistics for factory f1

```{r}
  
spc_data_f1<-dQ3_expected_f1 %>% 
  select(-factory) %>% 
  group_by(timepoint,week) %>% 
  summarise(xbar=mean(value),
         Range=max(value)-min(value)) 

spc_data_f1_table<-spc_data_f1 %>% 
  ungroup() %>% 
  slice_max(n=10, order_by = week) %>% 
  kbl() %>%
  kable_styling()
  

number_row_spc_data_f1<-spc_data_f1 %>% 
  nrow()

spc_data_f1_average_xbar_range <- spc_data_f1 %>%
  ungroup() %>% #group_by(week) %>%
  summarise(Average_xbar = mean(xbar),
         Average_Range = mean(Range))
   


```

Statistics table for spc_data_f1 is:  `r spc_data_f1_table`.
There are  `r number_row_spc_data_f1` number  of rows in spc_data_f1.# add extra line
Average across xbar and Range is:  `r spc_data_f1_average_xbar_range`

# Part5: Build Shewhart control charts for factory f1

```{r}

XbarR_constants_week_n5 <- read_csv("Data/XbarR_constants.csv") %>% 
  filter(n==5)

dQ5 <- spc_data_f1_average_xbar_range  %>%
  mutate(
    LCL = Average_Range * XbarR_constants_week_n5$D3,
    UCL = Average_Range * XbarR_constants_week_n5$D4
  )

ggplot(data=spc_data_f1, aes(x=timepoint)) +
geom_line(color="red", aes(x=timepoint, y=dQ5$LCL))+
geom_line(color="red", aes(x=timepoint, y=dQ5$UCL))+
geom_line(data=spc_data_f1, color="black", aes(y=Range),linetype="dotted")+
  labs(title = "Lineplot of Range vs Timepoint for factory f1", 
       subtitle = "LCL and UCL values in red and average range value in black",
       y="Range values",x="Timepoint") 


dQ5_xbar<-spc_data_f1_average_xbar_range   %>% 
  mutate(LCL=Average_xbar-Average_Range*XbarR_constants_week_n5$A2,
         UCL=Average_xbar+Average_Range*XbarR_constants_week_n5$A2)

ggplot(data=spc_data_f1, aes(x=timepoint)) +
  geom_line(color="black", aes(y=xbar),linetype="dotted")+
  geom_line(color="red", aes(x=timepoint, y=dQ5_xbar$LCL))+
  geom_line(color="red", aes(x=timepoint, y=dQ5_xbar$UCL))+
  labs(title = "Lineplot of Average xbar vs Timepoint for factory f1", 
       subtitle = "LCL and UCL values in red and average xbar value in black",
       y="Xbar value",x="Timepoint") + 
  ggthemes::theme_economist_white(base_size = 11, base_family = "sans",
  gray_bg = TRUE, horizontal = TRUE)

```

# Part6: Highlight special causes of variations

```{r}

vec<-spc_data_f1 %>% 
  pull(xbar)

mean_observed<-mean(vec)

highlight<-spc_data_f1 %>% 
  ungroup() %>% 
  mutate(is_between=between(xbar,dQ5_xbar$LCL,dQ5_xbar$UCL),
         slide_above_below=case_when(
           xbar>=mean_observed ~ 1,
           xbar<mean_observed ~ -1
         ),
       moving_ten_sum= abs(slide_sum(x = slide_above_below,  before = 10L, step=1L)),
       highlight_true=if_else(moving_ten_sum>10, TRUE, FALSE), 
       colour_def= case_when(
         highlight_true==TRUE ~ 'blue',
         xbar=!between(xbar, dQ5_xbar$LCL, dQ5_xbar$UCL) ~ 'blue',
        TRUE ~ 'black'
       ))
         # roll_sum = RcppRoll::roll_sum(slide_above_below, 10, align = "right")

# create general highlight =
# colors differ depending whether highlighting the point or not

# unable to change colours

highlight %>% 
ggplot(aes(x=timepoint, y=xbar)) +
  geom_line(aes(y=xbar ), linetype="dotted")+
  geom_point(aes(y=xbar, color=colour_def))+
  geom_line(color="black", aes(x=timepoint, y=dQ5_xbar$LCL))+
  geom_line(color="black", aes(x=timepoint, y=dQ5_xbar$UCL))+
  labs(title = "Lineplot of Average xbar vs Timepoint for factory f1", 
       subtitle = "LCL and UCL values in black, average xbar value in red, outliers in blue",
       y="Xbar value",x="Timepoint") + 
  ggthemes::theme_economist_white(base_size = 11, base_family = "sans",
  gray_bg = TRUE, horizontal = TRUE)+
  theme(legend.position = "none" )

           


```


# Part7: Build Shewhart control charts for factory f2

```{r}

spc_data_f2 <- dQ3_expected_f2 %>%
  select(-factory) %>%
  group_by(timepoint, week) %>%
  summarise(xbar = mean(value),
            Range = max(value) - min(value)) %>%
  ungroup()

spc_data_f2_average_xbar_range <- spc_data_f2 %>%
  summarise(Average_xbar = mean(xbar),
         Average_Range = mean(Range))


dQ7<-spc_data_f2_average_xbar_range  %>% 
  mutate(LCL=Average_Range*XbarR_constants_week_n5$D3,
         UCL=Average_Range*XbarR_constants_week_n5$D4)


vec_dQ7<-spc_data_f2 %>% 
  pull(xbar)

mean_observed_dQ7<-mean(vec)

highlight_dQ7<-spc_data_f2 %>% 
  ungroup() %>% 
  mutate(is_between=between(xbar,dQ7_xbar$LCL,dQ7_xbar$UCL),
         slide_above_below=case_when(
           xbar>=mean_observed ~ 1,
           xbar<mean_observed ~ -1
         ),
       moving_ten_sum= abs(slide_sum(x = slide_above_below,  before = 10L, step=1L)),
       highlight_true=if_else(moving_ten_sum>10, TRUE, FALSE), 
       colour_def= case_when(
         highlight_true==TRUE ~ 'blue',
         xbar=!between(xbar, dQ7_xbar$LCL, dQ7_xbar$UCL) ~ 'blue',
        TRUE ~ 'black'
       ))

highlight_dQ7 %>% 
ggplot(aes(x=timepoint, y=xbar)) +
  geom_line(aes(y=xbar ), linetype="dotted")+
  geom_point(aes(y=xbar, color=colour_def))+
  geom_line(color="black", aes(x=timepoint, y=dQ7_xbar$LCL))+
  geom_line(color="black", aes(x=timepoint, y=dQ7_xbar$UCL))+
  labs(title = "Lineplot of Average xbar vs Timepoint for factory f1", 
       subtitle = "LCL and UCL values in black, average xbar value in red, outliers in blue",
       y="Xbar value",x="Timepoint") + 
  ggthemes::theme_economist_white(base_size = 11, base_family = "sans",
  gray_bg = TRUE, horizontal = TRUE)+
  theme(legend.position = "none" )



ggplot(data = spc_data_f2, aes(x = timepoint)) +
  geom_line(color = "red", aes(x = timepoint, y = dQ7$LCL)) +
  geom_line(color = "red", aes(x = timepoint, y = dQ7$UCL)) +
  geom_line(data = spc_data_f2,
            color = "black",
            aes(y = Range),
            linetype = "dotted")+
labs(
  title = "Lineplot of Range vs Timepoint for factory f2",
  subtitle = "LCL and UCL values in red and average range value in black",
  y = "Range values",
  x = "Timepoint"
) 


dQ7_xbar<-spc_data_f2_average_xbar_range   %>% 
  mutate(LCL=Average_xbar-Average_Range*XbarR_constants_week_n5$A2,
         UCL=Average_xbar+Average_Range*XbarR_constants_week_n5$A2)

ggplot(data=spc_data_f2, aes(x=timepoint)) +
  geom_line(color="black", aes(y=xbar),linetype="dotted")+
  geom_line(color="red", aes(y=dQ7_xbar$LCL))+
  geom_line(color="red", aes(y=dQ7_xbar$UCL))+
  labs(title = "Lineplot of Average xbar vs Timepoint for factory f2", 
       subtitle = "LCL (red), UCL values (blue) and average xbar value in black",
       y="Xbar value",x="Timepoint") 

library(ggpomological)

  ggplot(data = spc_data_f1, aes(x= week, y= xbar))+
    geom_point() +
  labs(title = "Scatterplot of xbar vs week for factory f1", 
       y="Xbar value",x="week") + theme_pomological_fancy()+
    scale_color_pomological()
  
  ggplot(data = spc_data_f2, aes(x= week, y= xbar))+
    geom_point() +
  labs(title = "Scatterplot of xbar vs week for factory f2", 
       y="Xbar value",x="week") + ggthemes::theme_fivethirtyeight()
  

```
# Part8: Redefining montoring data for factory f2

```{r}

pacman::p_load(
  tidyverse,      # for data management and viz
  slider,         # for calculating moving averages
  tidyquant       # for calculating moving averages within ggplot
)

dQ8<-spc_data_f2 %>%
  group_by(week) %>%
  summarise(Average_Range=mean(Range, na.rm=T)) %>% 
   mutate(                                
    MR_bar = slide_dbl(
      Average_Range,                        
      .f = ~mean(.x, na.rm = T),         
      .before = 1))
    
  dQ8<-dQ8 %>% 
    mutate(Average_Average_Range=mean(Average_Range),
           LCL= Average_Average_Range-2.66*MR_bar,
         UCL=  Average_Average_Range+2.66*MR_bar)

ggplot(data=dQ8, aes(x=week)) +
  geom_line(color="black", aes(y=Average_Range),linetype="dotted")+
  geom_line(color="red", aes(y=LCL))+
  geom_line(color="red", aes(y=UCL))+
  labs(title = "Lineplot of Range vs Timepoint (by week) for factory f2", 
       subtitle = "LCL and UCL values in red and average range value in black",
       y="Range values",x="week") 

dQ8_xbar<-spc_data_f2 %>%
  group_by(week) %>%
  summarise(Average_xbar=mean(xbar, na.rm=T)) %>% 
   mutate(                                
    MR_bar = slide_dbl(
     Average_xbar,                        
      .f = ~mean(.x, na.rm = T),         
      .before = 1))

dQ8_xbar<-dQ8_xbar %>% 
    mutate(Average_Average_xbar=mean(Average_xbar),
           LCL= Average_Average_xbar-2.66*MR_bar,
         UCL=  Average_Average_xbar+2.66*MR_bar)

ggplot(data=dQ8_xbar, aes(x=week)) +
  geom_line(color="black", aes(y=Average_xbar),linetype="dotted")+
  geom_line(color="red", aes(y=LCL))+
  geom_line(color="red", aes(y=UCL))+
  labs(title = "Lineplot of xbar  vs Timepoint (by week) for factory f2", 
       subtitle = "LCL and UCL values in red and average range value in black",
       y="Range values",x="week") 

```
