---
title: "Project 4"
output: github_document
---

```{r}

library(tidyr)
library(readxl)
library(stringr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(knitr)

athletes_data <- read_excel("~/Dropbox/Courses/EPFL Extension School/Rstudio course/Final Projects/Github/Final_submission/20200817_Git_repository/project-fds-c1-s26-3561-1887/Report_4/olympics.xlsx", 
    sheet = "athletes")
country_data<-read_excel("~/Dropbox/Courses/EPFL Extension School/Rstudio course/Final Projects/Github/Final_submission/20200817_Git_repository/project-fds-c1-s26-3561-1887/Report_4/olympics.xlsx",
    sheet = "country")
games_data<-read_excel("~/Dropbox/Courses/EPFL Extension School/Rstudio course/Final Projects/Github/Final_submission/20200817_Git_repository/project-fds-c1-s26-3561-1887/Report_4/olympics.xlsx", 
    sheet = "games")
medals_data<-read_excel("~/Dropbox/Courses/EPFL Extension School/Rstudio course/Final Projects/Github/Final_submission/20200817_Git_repository/project-fds-c1-s26-3561-1887/Report_4/olympics.xlsx", 
    sheet = "medals")
```

# Part 1

some athletes competed for different countries over time

```{r echo=TRUE, message=FALSE, warning=FALSE}

athlete_participated_different_country<-country_data %>% 
  select(-Age,-Games)%>%
  group_by(athlete_id)%>%
  mutate(participation=n()) %>% 
  filter(participation!=1) %>% 
  count()

athlete_participated_different_country_table<-athlete_participated_different_country %>% 
  top_n(20) %>% 
  knitr::kable()

```

Yes, many athletes (37121) have participated for different countries.
    
# Part 2

ten athletes that took part in most games

```{r echo=TRUE, message=FALSE, warning=FALSE}
Participation_number_of_games<-medals_data%>% 
  select(athlete_id,Games) %>% 
  group_by(athlete_id) %>% 
  summarise(number_of_games=n())%>%
  arrange(desc(number_of_games))%>%
  top_n(10)

Participation_number_of_games_table<-Participation_number_of_games %>% 
  kable()

```

ten athletes that took part in most games are :`r Participation_number_of_games_table`

# Part 3

athlete(s) kept a Gold medal for the longest time

```{r echo=TRUE, message=FALSE, warning=FALSE}
Athlete_gold_longest_time<-medals_data%>% 
  select(athlete_id,Games,Sport,Medal)%>% 
  mutate(Year=as.numeric(str_sub(Games,start = 1L,end = 4L)))%>% 
  filter(Medal=="Gold")%>%
  group_by(athlete_id)%>%
  mutate(number_of_years=max(Year)-min(Year))%>%
  arrange(desc(number_of_years)) 
  
  Athlete_gold_longest_time_by_sport<-Athlete_gold_longest_time %>%
    select(Sport,number_of_years) %>% 
    group_by(Sport) %>% 
    top_n(1)

Athlete_gold_longest_time_table<-Athlete_gold_longest_time %>% 
  kable()
```

Fencing athlete(s) kept a Gold medal for the longest time.

# Part 4

country(ies) kept a Gold medal for the longest time

```{r echo=TRUE, message=FALSE, warning=FALSE}

Countries_gold_longest_time<-medals_data%>% 
  select(athlete_id,Games,Team,Sport,Medal)%>% 
  mutate(Year=as.numeric(str_sub(Games,start = 1L,end = 4L)))%>% 
  filter(Medal=="Gold")%>%
  group_by(Team,Sport)%>%
  mutate(number_of_years=max(Year)-min(Year))%>%
  ungroup(Sport) %>% 
  arrange(desc(number_of_years)) 


Countries_gold_longest_time_table<-Countries_gold_longest_time %>% 
  select(Team,Sport,number_of_years) %>% 
  filter(number_of_years==120) %>% 
  group_by(Team) %>% 
  count() %>% 
  kable()

```

country(ies) kept a Gold medal for the longest time are :`r Countries_gold_longest_time_table`

# Part 5

ten athletes that competed in the most events (some athletes take part in more than one event during games)

```{r echo=TRUE, message=FALSE, warning=FALSE}
Athletes_participated_most_events<-medals_data%>% 
  select(athlete_id,Event)%>%
  group_by(athlete_id) %>% 
  summarise(number_of_events=n())%>%
  arrange(desc(number_of_events))%>%
  top_n(10)   #Not sure why answer for Part2 and Part 5 are same???

Athletes_participated_most_events_table<-Athletes_participated_most_events %>% 
  kable()
```

ten athletes that competed in the most events :`r Athletes_participated_most_events_table`

# Part 6

number of medals per country (rows) and per year (column)

```{r echo=TRUE, message=FALSE, warning=FALSE}

Number_of_medals_per_country<-medals_data%>% 
  select(Games,Team,Medal)%>% 
  mutate(Year=as.numeric(str_sub(Games,start = 1L,end = 4L)))%>% 
  select(-Games)%>% 
  group_by(Team,Year)%>% 
  summarise(total_number_of_medals_per_country=n())%>%
  ungroup() %>%
  top_n(15)

Number_of_medals_per_country_table<-Number_of_medals_per_country %>% 
  kable()
```

number of medals per country (rows) and per year (column) are :`r Number_of_medals_per_country_table`


# Part 7

Relationship between country and the probability of winning a medal

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Sorry I did not realize till recently that i did not finish Part 7, will try my best before the project defense...

#Total_number_of_medals_counted_by_year<-Number_of_medals_per_country %>% 
#group_by(Year) %>% 
#summarise(Total_number_of_medals_counted=sum(total_number_of_medals_per_country))
```
  
  
# Part 8

Scatterplot showing the average height and weight of competitors per sport

```{r echo=TRUE, message=FALSE, warning=FALSE}
medals_atheltes_left_join<-medals_data%>%
  left_join(athletes_data, 
            by = c("athlete_id" ="ID"))%>%
  mutate(BMI=Weight/(Height/100)^2)

Average_weight_by_sport<-medals_atheltes_left_join%>%
  group_by(Sport)%>%
  summarize(mean_weight = mean(Weight, na.rm = TRUE))
Average_height_by_sport<-medals_atheltes_left_join%>%
  group_by(Sport)%>%
  summarize(mean_height = mean(Height, na.rm = TRUE))

Average_BMI_by_sport<-medals_atheltes_left_join%>%
  group_by(Sport)%>%
  summarize(mean_BMI = mean(BMI, na.rm = TRUE))

largest_weight_by_sport <- Average_weight_by_sport %>%
  slice(which.max(mean_weight))
smallest_weight_by_sport <- Average_weight_by_sport %>%
  slice(which.min(mean_weight)) 
largest_height_by_sport <- Average_height_by_sport %>%
  slice(which.max(mean_height))
smallest_height_by_sport <- Average_height_by_sport %>%
  slice(which.min(mean_height)) 
largest_BMI_by_sport <- Average_BMI_by_sport %>%
  slice(which.max(mean_BMI))
smallest_BMI_by_sport <- Average_BMI_by_sport %>%
  slice(which.min(mean_BMI)) 

weight_height<-left_join(Average_weight_by_sport,Average_height_by_sport, by=c("Sport"))
weight_height_BMI<-left_join(weight_height,Average_BMI_by_sport, by=c("Sport"))

ggplot2::ggplot(data= weight_height_BMI,
                aes(y=mean_weight, 
                    x=mean_height,
                    color=Sport)) + geom_point() +
  theme(legend.position="none")+
  geom_text(data= weight_height_BMI%>%
              slice(which.max(mean_weight)),
            aes(label=stringr::str_c(largest_weight_by_sport$Sport,
                                     "largest weight by sport",sep="-")), check_overlap = T) +
  geom_text(data= weight_height_BMI%>%
              slice(which.min(mean_weight)),
            aes(label=stringr::str_c(smallest_weight_by_sport$Sport,
                                     "smallest weight by sport",sep="-")), check_overlap = T) +
  
  geom_text(data= weight_height_BMI%>%
              slice(which.max(mean_height)),
            aes(label=stringr::str_c(largest_height_by_sport$Sport,
                                     "largest height by sport",sep="-")), check_overlap = T) +
  
  geom_text(data= weight_height_BMI%>%
              slice(which.min(mean_height)),
            aes(label=stringr::str_c(smallest_height_by_sport$Sport,
                                     "smallest height by sport",sep="-")), check_overlap = T) +
  
  geom_text(data= weight_height_BMI%>%
              slice(which.max(mean_BMI)),
            aes(label=stringr::str_c(largest_BMI_by_sport$Sport,
                                     "largest BMI by sport",sep="-")), check_overlap = T) +
  
  geom_text(data= weight_height_BMI%>%
              slice(which.min(mean_BMI)),
            aes(label=stringr::str_c(smallest_BMI_by_sport$Sport,
                                     "smallest BMI by sport",sep="-")), check_overlap = T) +
  
  labs(title = "average height and weight of competitors per sport",
       subtitle = "Using colors to differentiate sport wise average weight and height of participants",
       caption = "Source: Olympics",
       x = "Average height",
       y = "Average weight",
       fill = "Sport")
```

# Part 9 

Number of medals (gold, silver and bronze given per year, facet chart: summer and winter)

```{r echo=TRUE, message=FALSE, warning=FALSE} 
Medals_category_year_session<-medals_data%>% 
  select(Games,Medal)%>% 
  mutate(Year=as.numeric(str_sub(Games,start = 1L,end = 4L)))%>%
  mutate(Session=str_sub(Games,start = 5L))%>%
  select(-Games)%>%
  arrange(desc(Year))

medals_per_year <- Medals_category_year_session%>%
  group_by(Year, Medal,Session)%>%
  summarise(number_of_medals=n())%>%
  ungroup()

# Line plot for Part 9

medals_per_year %>% drop_na() %>% 
  ggplot(mapping=aes(x=Year, 
                     y=number_of_medals)) +
  geom_line(aes(group=Medal, color=Medal)) +
  theme(legend.position="bottom", legend.box = "horizontal")+ 
  facet_wrap(vars(Session))+
  labs (title = "Evolution of total number of medals per year",
        subtitle = "Using colors to differentiate medal categories",
        caption = "Source: Olympics",
        x = "Year",
        y = "Total number of medals")
```



